### 2010-03-29 papers/icfp-2010-desc/paper_desc.tex
Desc@ICFP: precise that Desc is not here to rivalize with Generic Haskell or related. It is just pedagogical.
Desc@ICFP: use 'pattern functor' instead of 'signature functor', which seems to be more used in the Generic prog. literature.

In this section, we describe an implementation of inductive types, as
we know them in simply-typed languages such as Haskell or OCaml. This
choice is motivated by pedagogical considerations: by describing the
layman data-types in our setting, we build on our intuition of
data-types we are familiar with. However, our goal is and remains to
discuss data-types as we find them in dependently-typed languages.

@@

Our first example is the natural numbers, or rather its signature
functor. Our code is presented in the high-level expression language
of Section~\ref{sec:type-propagation}. The translation back to the raw
terms is laborious but should not pose any difficulty. The code is the
following:

@@

With a small change to the definition of $\NatD$, we obtain the
signature of lists:

@@

So far, we have used our universe of descriptions to build
functors. To illustrate its usage, we have implemented the signature
functors of natural numbers, lists, and binary trees. However, the
class of functors expressible with descriptions enjoy another
property: they all admit a fix-point. Indeed, descriptions are
interpreted into \emph{polynomial functors}~\cite{who?}\note{who is
  the relevant source for polynomial functors?}, as they are only
composed of $\Sigma$-types, the unit type, and variables.

@@

Whereas simply-typed languages feature a convenient grammar for
defining data-types and manipulating them, our proposal could be seen
as a step backward. In our setting, the user would have to write a
description instead of a sugared data-type definition. Then, to
recognise an element of $\Nat$ for instance, she would have to
identify $\Mu{}$ of a description as indeed the fix-point of the
functor signature of natural numbers. Finally, she would have to write
codes instead of the natural data-type constructor.

@@

Doing so, we seal data-type definition behind a convenient
abstraction. We take advantage of this abstraction barrier to tamper
the second issue. Namely, our user would rather be presented a name
rather than a fix-point of a signature functor. Surprisingly, it is
sufficient to \emph{label} the $\Mu{}$ constructor with the
user-provided name. When pretty-printing this object, we just expose
the label, instead of the definition in full.

>>>>

In this section, we describe an implementation of inductive types, as
we know them in simply-typed languages such as Haskell or OCaml. This
choice is motivated by pedagogical considerations: by describing the
layman data-types in our setting, we build on our intuition of
data-types we are familiar with. In particular, we do not aim at
describing a model for generics in simply-typed functional
language. Our goal is and remains to discuss data-types as we find
them in dependently-typed languages.

@@

Our first example is the natural numbers, or rather its pattern
functor. Our code is presented in the high-level expression language
of Section~\ref{sec:type-propagation}. The translation back to the raw
terms is laborious but should not pose any difficulty. The code is the
following:

@@

With a small change to the definition of $\NatD$, we obtain the
pattern functor for lists:

@@

So far, we have used our universe of descriptions to build
functors. To illustrate its usage, we have implemented the pattern
functors of natural numbers, lists, and binary trees. However, the
class of functors expressible with descriptions enjoy another
property: they all admit a fix-point. Indeed, descriptions are
interpreted into \emph{polynomial functors}~\cite{who?}\note{who is
  the relevant source for polynomial functors?}, as they are only
composed of $\Sigma$-types, the unit type, and variables.

@@

Whereas simply-typed languages feature a convenient grammar for
defining data-types and manipulating them, our proposal could be seen
as a step backward. In our setting, the user would have to write a
description instead of a sugared data-type definition. Then, to
recognise an element of $\Nat$ for instance, she would have to
identify $\Mu{}$ of a description as indeed the fix-point of the
pattern functor of natural numbers. Finally, she would have to write
codes instead of the natural data-type constructor.

@@

Doing so, we seal data-type definition behind a convenient
abstraction. We take advantage of this abstraction barrier to tamper
the second issue. Namely, our user would rather be presented a name
rather than a fix-point of a pattern functor. Surprisingly, it is
sufficient to \emph{label} the $\Mu{}$ constructor with the
user-provided name. When pretty-printing this object, we just expose
the label, instead of the definition in full.

### 2010-03-29 papers/icfp-2010-desc/paper_desc_levitation.tex
Desc@ICFP: use 'pattern functor' instead of 'signature functor', which seems to be more used in the Generic prog. literature.

Let us examine the consequences of our act. First of all, we discover
that the type theory does not actually need to be extended with the
type former $\EnumU$, the constructors $\NilE$ and $\ConsE$. We have
just defined $\EnumU$ above, by fix-point over the signature functor denoted
by $\ListD$. $\NilE$ and $\ConsE$ are simply the $\ListNil$ and $\ListCons$
constructors of lists.

@@

Being an inductive type, this $\FMFreeMonad$ data-type is itself
defined by a functor signature. It is given by:

>>>>

Let us examine the consequences of our act. First of all, we discover
that the type theory does not actually need to be extended with the
type former $\EnumU$, the constructors $\NilE$ and $\ConsE$. We have
just defined $\EnumU$ above, by fix-point over the pattern functor denoted
by $\ListD$. $\NilE$ and $\ConsE$ are simply the $\ListNil$ and $\ListCons$
constructors of lists.

@@

Being an inductive type, this $\FMFreeMonad$ data-type is itself
defined by a pattern functor. It is given by:

### 2010-03-29 papers/icfp-2010-desc/paper_idesc.tex
Desc@ICFP: use 'pattern functor' instead of 'signature functor', which seems to be more used in the Generic prog. literature.

In order to gain some intuition with $\IDesc$, let us re-implement the
signature functor of natural numbers. We refer the reader to
Section~\ref{sec:desc-examples} for its incarnation in the universe of
descriptions. The code is the following:

@@

Another instance of free monad is $\IDesc$ itself. Indeed, $\DVar$ is
nothing but the \return. The remaining constructors are simply the
signature functor, trivially indexed by $\Unit$. Hence, we describe
this signature functor by the following code:

>>>>

In order to gain some intuition with $\IDesc$, let us re-implement the
pattern functor of natural numbers. We refer the reader to
Section~\ref{sec:desc-examples} for its incarnation in the universe of
descriptions. The code is the following:

@@

Another instance of free monad is $\IDesc$ itself. Indeed, $\DVar$ is
nothing but the \return. The remaining constructors are simply the
pattern functor, trivially indexed by $\Unit$. Hence, we describe
this functor by the following code:
